---
title: "Multiple Responses"
author: "Taylor Pospisil"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple Responses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

RFCDE extends to multiple response variables to estimate joint
conditional densities. Consider data from the following model

\begin{eqnarray}
X &\sim& U(0, 1) \\
Z_{1} &\sim& U(0, X) \\
Z_{2} &\sim& U(X, Z_{1})
\end{eqnarray}

This exhibits dependence not only between the responses and the
covariates, but also between the responses themselves. Thus merely
estimating the univariate conditional densities would lose the
information about the dependence between $Z_{1}$ and $Z_{2}$.

```{r}
library(RFCDE)
set.seed(42)

generate_data <- function(n) {
  x <- runif(n)
  z1 <- runif(n, 0, x)
  z2 <- runif(n, z1, x)

  return(list(x = x, z = cbind(z1, z2)))
}

n_train <- 10000

train_data <- generate_data(n_train)
x_train <- train_data$x
z_train <- train_data$z
```

## Training

Training is the same as for the univariate case. A tensor basis is
used for the density estimates at each node; you can specify the
number of basis functions for each dimension or just specify a single
number which is applied to every dimension. Note that this scales like
$n^d$ for $n$ basis functions in $d$ dimensions.

```{r}
n_trees <- 100
mtry <- 1
node_size <- 20
n_basis <- c(15, 15)

forest <- RFCDE(x_train, z_train, n_trees = n_trees, mtry = mtry,
                node_size = node_size, n_basis = n_basis)
```

## Prediction

Prediction requires a multivariate grid of points at which to evaluate
the density. Bandwidths can be specified by a covariance matrix, a
float (which scales the identity covariance) or a string (see
documentation for details).

```{r}
n_grid <- 30
z_grid <- expand.grid(z1 = seq(0, 1, length.out = n_grid),
                      z2 = seq(0, 1, length.out = n_grid))
density <- predict(forest, c(0.3), "CDE", z_grid, bandwidth = "auto")

image(seq(0, 1, length.out = n_grid), xlab = "Z1",
      seq(0, 1, length.out = n_grid), ylab = "Z2",
      z = matrix(density, n_grid, n_grid))
```