---
title: "RFCDE"
author: "Taylor Pospisil"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RFCDE Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Random forests is a common non-parametric regression technique which
performs well for mixed-type data and irrelevant covariates, while
being robust to monotonic variable transformations. RFCDE fits random
forest models optimized for nonparametric conditional density
estimation.

## Example

```{r}
library(RFCDE)
set.seed(42)

generate_data <- function(n) {
  x_relevant <- matrix(runif(n * 10), n, 10)
  x_irrelevant <- matrix(runif(n * 10), n, 10)
  z <- rnorm(n, rowSums(x_relevant), 1)
  return(list(x = cbind(x_relevant, x_irrelevant), z = z))
}

n_train <- 10000
n_test <- 4

train_data <- generate_data(n_train)
x_train <- train_data$x
z_train <- train_data$z
```

### Training

Trees are recursively partitioned to minimize the CDE loss

$$ \int \int (\hat{f}(x \mid z) - f(x \mid z))^{2} dz dP(x) $$

This is efficiently calculated using an orthogonal series
representation of the conditional densities. The resolution of this
representation is controlled by `n_basis`.

```{r}
n_trees <- 1000
mtry <- 4
node_size <- 20
n_basis <- 15

forest <- RFCDE(x_train, z_train, n_trees = n_trees, mtry = mtry,
                node_size = node_size, n_basis = n_basis)
```

### Prediction

We use the forest structure to determine weights for a weighted kernel
density estimate. The `predict` function evaluates this density on the
provided grid.

```{r}
bandwidth <- 0.2
n_grid <- 100
z_grid <- seq(0, 10, length.out = n_grid)
x_test <- generate_data(1)$x

density <- predict(forest, x_test, "CDE", z_grid, bandwidth = bandwidth)
```

```{r}
plot(z_grid, density, type = "l", col = "green")
lines(z_grid, dnorm(z_grid, sum(x_test[1:10]), 1))
```